---
title: "get_ready_model"
author: "Zach"
date: "2023-12-14"
output: html_document
---

```{r}
# * load packages ----
rm(list = ls())
library(devtools)
library(dplyr)
library(gganimate)
library(ggforce)
library(ggplot2)
library(readr)
```

This is the tracking data from week 1
```{r}
tracking <- read.csv('tracking_with_tacklesDec13.csv')
```

This is the tackle data from week1
```{r}
tackles <- read.csv('tackles.csv')
```

Ok lets mark who made first contact on a play
```{r}
frameId_first_contact <- tracking %>% filter(event == 'first_contact') %>% pull(game_play_frameId)
frameId_first_contact <- unique(frameId_first_contact)
#get frameIds all the frames that are listed as "first contact" 
```

```{r}
tracking %>% filter(on_defense == 1, is.na(dist_from_bc))
  
```


```{r}
frame_table <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(frame_table) <- c('game_playId', 'nflId')
counter = 0
for (frame in frameId_first_contact){
  counter= counter + 1
  print(counter)
  #for every frame that is first contact
  d_play_data <- tracking %>% filter(game_play_frameId == frame, on_defense == 1)
  #get tracking data of frame of just defenseive players
  dis_closest_bc <- min(d_play_data$dist_from_bc)
  #closest distance to bc
  row_to_add <- d_play_data %>% filter(dist_from_bc == dis_closest_bc)
  
  frame_table <-rbind(frame_table, row_to_add)
  #add to running table
}
#every row in this data set made first contact
#set marker to 1
frame_table$made_first_contact <- 1
#merge this with tracking
write.csv(frame_table, "first_contact.csv", row.names=FALSE)
```

```{r}
frame_table %>% ggplot() + geom_histogram(aes(x= dist_from_bc))
```
Ok this shows how far the tackler is when he makes frist contact. 

```{r}
frame_table %>% ggplot() + geom_boxplot(aes(x= dist_from_bc))
```
```{r}
summary(frame_table$dist_from_bc)
```

These outliers are likely errors in the data such as stating "first_contact" in the wrong place. I will filter out anything above 2 and say that player did not make first contact. 

```{r}
frame_table <- frame_table %>% filter(dist_from_bc<2) 
```

```{r}
frame_table %>% ggplot() + geom_boxplot(aes(x = dist_from_bc))
```
```{r}
summary(frame_table$dist_from_bc)
```

```{r}
frame_table <- read.csv('first_contact.csv')
```

```{r}
frame_table %>% select(c('game_playId','made_first_contact' )) %>% merge(tracking, all = TRUE)
```

```{r}
write.csv(tracking, "tracking_with_tacklesDec13.csv", row.names=FALSE)
```

The next step is to define the moment the tackler makes initial contact with the ball carrier. PFF refrence has spotters manually input when first contact of the play occurs. This does introduce human error, but there are strict guidlines on how to define first contact. Above states the player that makes initial contact at median distance of .88 away from ball carrier. We will use this median to find the point that each tackler makes initial contact. This will be the frame at which the tackler is closest to .88


First we need to add in who made tackles 

```{r}

tackle_df <- tackles %>% filter(gameId  %in% unique(tracking$gameId), assist == 0, forcedFumble == 0) %>%select('gameId', 'playId', 'nflId', 'tackle', 'pff_missedTackle')

tackle_df$tackle_opp =1
```


```{r}
tackle_df %>% select(tackle) %>% table() %>% prop.table()
```

Looks like we have 17% missed tackles and 83% made tackles 

Time to merge with tracking data




```{r}
#tracking <- tackle_df %>% merge(tracking, all = TRUE)
write.csv(tracking, "tracking_with_tacklesDec13.csv", row.names=FALSE)
```
The number of observations did not change, so this merged correctly! 

Now we want to find the frames that initial contact is made. For the `tackle_opps` that made first contact this is easy, but for the rest we will use the median value

```{r}
tracking
```




```{r}
frame_of_tackle_opp  <- data.frame(matrix(ncol = 30, nrow = 0))
colnames(frame_of_tackle_opp) <- colnames(tracking)
counter = 0 
id = 0
errors <- c()

for (play in unique(tracking$game_playId )){
  #for each game id combo
  counter = counter + 1
  print(counter)
  id = 1+ id
  play_data <- tracking %>% filter(game_playId == play)
  #each play data 
  list_of_tacklers <- play_data %>% filter(tackle_opp ==1) %>% pull(nflId)
  # get all of the players that had a tackle oppurtunity in the play
  if (length(list_of_tacklers)>0){
    #if there is a tackler in play
    list_of_tacklers <- unique(list_of_tacklers)
    for (tacklerId in list_of_tacklers){
      id = id+1
      #for each tackler
      tackler_data <- play_data %>% filter(nflId == tacklerId)
      #tackler data
      if (!('out_of_bounds' %in% unique(tackler_data$event))){
      if (1 %in% tracking$made_first_contact){
        #if made first contact is set to 1 this tackler made first contact and that frame will be what we use for tackler_contact_frame
        tackler_contact_frame <- tackler_data%>% filter(event == 'first_contact') %>% pull(frameId) }
        
      else{
        #if player is not marked for initial contact we need to approximate
        tackler_data$dev_median <- abs(tackler_data$dist_from_bc - .88)
        tackler_contact_frame <- tackler_data %>% filter( dev_median == min(tackler_data$dev_median)) %>% pull(frameId)}
      
        tackler_frame_data <- play_data %>% filter(frameId == (tackler_contact_frame -1))
        
        if (length(tackler_frame_data$nflId)>0){
        tackler_frame_data$tackle_oppId <- id
        #the 3 lines below set all tackle data to NA if they are not the tackler         in frame.
        tackler_frame_data[tackler_frame_data$nflId != tacklerId,]$tackle = NA
        tackler_frame_data[tackler_frame_data$nflId != tacklerId,]$pff_missedTackle = NA
        tackler_frame_data[tackler_frame_data$nflId != tacklerId,]$tackle_opp = NA
        frame_of_tackle_opp <- rbind(frame_of_tackle_opp, tackler_frame_data)}
        else 
          errors <- append(errors, play)
        }
  }}}
  
  
  
  
```






```{r}
length(unique(frame_of_tackle_opp$tackle_oppId))
```
There are 1181 tackle oppurtunities that we will analyze 

```{r}
write.csv(frame_of_tackle_opp, "tackle_oppDec13.csv", row.names=FALSE)
```

Let look at the distance of our tacklers from ball one frame before contact to see if we did calculations right. 
```{r}
frame_of_tackle_opp %>% filter(tackle_opp ==1)%>% ggplot() + geom_histogram(aes(x = dist_from_bc))
```
Obviously we did some manual calculations, and once again there are probably some errors just within labeling, we will filter those out. 

```{r}
frame_of_tackle_opp %>% filter(tackle_opp ==1)%>% ggplot() + geom_boxplot(aes(x = dist_from_bc))
```
```{r}
frame_of_tackle_opp %>% filter(tackle_opp ==1)%>% select(dist_from_bc) %>% summary()
```
This is really encouraging that we did this properly! Players relatively close one frame before what we believe contact to be. 

```{r}
frame_of_tackle_opp %>% filter(tackle_opp ==1, dist_from_ball<3.5)%>% ggplot() + geom_boxplot(aes(x = dist_from_bc))
```
This removes a lot of our outliers. Will make this permanent, and potentially cut even closer later. 

I know this is not the most efficient, but there were a lot of issues with this step. So we are doing it the long way.Create a list of all the `tackl_oppId`s that are withing 3. 3 seems realistic distance from ball carrier to still make a tackle in the next frame.
```{r}
id_keep <- tacklers %>% filter(dist_from_bc <3, tackle_opp == 1) %>% pull(tackle_oppId)
```

Then filter for only those ids
```{r}
frame_of_tackle_opp <- frame_of_tackle_opp %>% filter(tackle_oppId %in% id_keep)
```

```{r}
frame_of_tackle_opp %>% filter(tackle_opp ==1)%>% ggplot() + geom_boxplot(aes(x = dist_from_bc))
```
This looks as expected. We now have a df of tracking data for the frame before a tackler makes initial contact

```{r}
write.csv(frame_of_tackle_opp, "tackle_oppDec13.csv", row.names=FALSE)
```



```{r}
df_tackle_opp <- read.csv('tackle_oppDec13.csv')
```



next we want a df of just the tackler and relevent statistics to predict if the tackle was made. Right now we will simply grab all of the tracking data. Along with the actual tackle stat


```{r}
df_to_model  <- data.frame(matrix(ncol = 18, nrow = 0))
colnames(df_to_model) <- colnames(c('tackle_oppId', 'tackle', 'pff_missedTackle', 'playDirection', 'x', 'y', 's', 'a', 'o', 'dir', 'made_first_contact','tackle_oppId', 'x_bc', 'y_bc', 's_bc', 'a_bc', 'o_bc', 'dir_bc' ))
counter = 0

for( id in unique(df_tackle_opp$tackle_oppId)) {
  counter = counter +1 
  print(counter)
  #for each tackle opp
  play_data <- df_tackle_opp %>% filter(tackle_oppId == id)
  #make df
  tackler_to_add <- play_data %>% filter(tackle_opp == 1) %>%
    select(tackle_oppId, tackle, pff_missedTackle, playDirection, x, y, s, a, o, dir, made_first_contact)
  #get tracking and tackle datafor tackler
  bc_to_add <- play_data %>% filter(has_ball == 1) %>% select(tackle_oppId, x, y, s, a, o, dir)
  #ball carrier tracking data
  colnames(bc_to_add) <- c('tackle_oppId', 'x_bc', 'y_bc', 's_bc', 'a_bc', 'o_bc', 'dir_bc')
  #rename to show this is ball carrier
  to_add <- merge(tackler_to_add, bc_to_add, all= TRUE)
  #merge the two
  df_to_model <- rbind(df_to_model, to_add)
  #add to df to model
}
```

```{r}
df_to_model[is.na(df_to_model$made_first_contact),]$made_first_contact = 0
```



```{r}
df_to_model
```

```{r}
write.csv(df_to_model, "df_to_model.csv", row.names=FALSE)
```


```{r}
tackle_oppdf <- read.csv('frame_of_tackle_opp.csv'
                         )
```


